---
title: "RNAfriends"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RNAfriends}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(RNAfriends)
```

## Package description

This package is based on the central dogma of molecular biology, which states that a DNA filament (sequence of nucleotides) is transcribed into RNA, which is finally translated into aminoacid sequences (that form proteins).

With this package, you can recreate a DNA sequence, transcribe it, translate it and visualize the relative abundance of the aminoacids.

The functions that this package contains are:

```{r}
RNAfriends::makeDNA()
RNAfriends::function_2()
RNAfriends::codonize()
RNAfriends::RNA_aminoacid_translation()
RNAfriends::plot_aminoacids_abundance()

```

## makeDNA

## function_2

## codonize

Splits a sequence of DNA or RNA into strings of codons.

```{r}
sequence = "ATGACATCGAUGTTAATGACATCGAUGTTAATGACATCGAUGTTAATCGAUGTTAATGACATCGAUGTTAATGACATCGAUGTTAATGAGACTTGACATCGAUGTAATCGAUGTTAATGACATTGACATCGAUGCGAUGTTAATGACATCGAUGTTAATGAGACATCGAUGTTAATGACATCGAUGTTAATGACATCGAUGTTAATGACATCGATCGAUGTTAATGACATCGAUGTTAATGACATCGAUGTTAATGACATCGAUGTTA"

RNAfriends::codonize(sequence, start = 1)
```

## RNA_aminoacid_translation

The function decodes a given RNA sequence to its aminoacid components.

```{r}
RNA_sequence = c("UUU", "AAA", "AUA")
peptide_sequence = RNAfriends::RNA_aminoacid_translation(RNA_sequence)

peptide_sequence
```

## plot_aminoacids_abundance

The function plots the aminoacids abundance for a given peptide-sequence.

```{r}
peptide_sequence = "QERRYQ"
RNAfriends::plot_aminoacids_abundance(peptide_sequence)
```

## Combined example
