---
title: "RNAfriends"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RNAfriends}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(RNAfriends)
```

## Package description

This package is based on the central dogma of molecular biology, which states that a DNA filament (sequence of nucleotides) is transcribed into RNA, which is finally translated into aminoacid sequences (that form proteins).

With this package, you can recreate a DNA sequence, transcribe it, translate it and visualize the relative abundance of the aminoacids.

The functions that this package contains are:

```{r}
RNAfriends::makeDNA()
RNAfriends::function_2()
RNAfriends::codonize()
RNAfriends::RNA_aminoacid_translation()
RNAfriends::plot_aminoacids_abundance()

```

## makeDNA

Use this function to generate a random sequence of DNA, of the length that you prefer. generateseq requires an integer that will define the length of your DNA sequence.

Bear in mind that, to obtain meaningful results with this package, your DNA length should be \>= 3.

Example:

```{r}
RNAfriends::makeDNA(250)
```

## function2

## codonize

## RNA_aminoacid_translation

## plot_aminoacids_abundance

## Combined example
